HUME: Human UCCA-Based Evaluation of Machine Translation

### 2016-03-18 paper/main.tex
section 2

\paragraph{Machine Translation Evaluation.}
%MT evaluation has been traditionally pursued through human-based
%evaluation that quantifies the adequacy of the translation in preserving the meaning 
%of the source sentence, and its fluency as a sentence in the target language \cite{lopez2008statistical}.
Human evaluation is generally done by ranking the outputs of multiple systems
e.g., in the WMT shared tasks \cite{bojar2015findings}, or by assigning
individual adequacy/fluency scores to each translation, a procedure recently improved
by \newcite{graham2015improving}.
However, while providing the gold standard for MT evaluation, human evaluation
is not a scalable solution.
%and is infeasible where a wide variety of systems or parameter settings are to be evaluated.
% has motivated the construction of cheaper methods that correlate with human
%judgement.

Scalability is addressed by employing automatic and semi-automatic approximations of human
judgements. Commonly, such scores decompose over the sub-parts of
the translation, and quantify how many of these sub-parts appear in a manually created reference translation.
This decomposition further allows system developers
to localize the errors, or at least unconfirmed parts within the translation.
The most commonly used measures decompose over n-grams or individual words, e.g., 
BLEU \cite{Papineni:2002}, NIST \cite{Doddington:2002} and METEOR  \cite{Banerjee:2005}.
Another common approach is to determine the similarity between the reference and translation
by applying string edit distance \cite{snover2006study}.
While these measures stimulated much progress in MT research by allowing
the evaluation of massive-scale experiments,
their decomposition over words and n-grams is better suited for localizing errors
in the output of shallow MT methods, and is less instructive in identifying the semantic
nature of the errors.

@@

In order to address this shortcoming, more recent work quantified
the similarity of the reference and translation in terms
of their structure. \newcite{liu2005syntactic} took a syntactic approach, 
quantifying the similarity between the syntactic structures of the reference and translation.
\newcite{owczarzak2007evaluating} took a similar approach using lexical-functional grammar structures.
\newcite{gimenez2007linguistic} proposed to aggregate multiple types of information,
capturing the overlap between the output and reference in terms of their
semantic (predicate-argument structures), lexical and morphosyntactic features.

@@

\begin{tabular}{|l|l|l|l|}
\hline
L & Linker &  A & Participant \\
H & Scene & R & Relator \\
P & Process & C & Centre \\
\hline
\end{tabular}
    \end{center}
\caption{\label{fig:ucca_example_v2}
  Sample UCCA annotation where leaves correspond to words and nodes correspond to units.
  Edge labels mark UCCA categories, which are described in the table.}
%which are largely ignored in this work
\end{figure}

\paragraph{Semantic Representation.}
UCCA (Universal Conceptual Cognitive Annotation, \inparcite{abend2013universal}) is a
cross-linguistically applicable, lightweight
scheme for semantic annotation. Formally, a UCCA structure is a directed acyclic graph (DAG),
whose leaves correspond to the words of the text.
The graph's nodes, called {\sc units}, are either terminals or several elements jointly
viewed as a single entity according to some semantic or cognitive consideration. Edges bear
a category, indicating the role of the sub-unit in the structure the unit
represents.

UCCA's current inventory of distinctions focuses on argument structures
(adjectival, nominal, verbal and others) and relations between them.
The most basic notion is the Scene, which describes a movement, an
action or a state which is persistent in time. Each Scene contains one main
relation and zero or more participants. For example, the sentence ``After graduation, Tom moved to NYC''
contains two Scenes, whose main relations are ``graduation'' and ``moved''.
The participant ``Tom'' is a part of both Scenes, while ``NYC'' only of the
latter (\figref{fig:ucca_example_v2}). Further categories account for
relations between Scenes and the internal structures of complex participants and relations.

% UCCA Motivation
UCCA is a natural candidate for defining a semantic
MT evaluation measure for a number of reasons. 
First, it defines an intuitive set of categories
that can be reliably annotated by non-experts, after as little as two hours
of training \cite{marinotti2014}.
Second, UCCA is a cross-linguistically applicable scheme, seeking to
represent what is shared between languages
by building on linguistic typological theory, and specifically on ``Basic Linguistic Theory''
\cite{Dixon:10a,Dixon:10b,Dixon:12}, one of the most widely used frameworks for linguistic description.
The cross-linguistic applicability of UCCA has been tested in annotations of
English, French, German and Czech.
Third, the scheme has been shown to be semantically stable
in translation: UCCA annotations of translated text usually contain the same set of relationships
\cite{sulem2015conceptual}. This is an indication that the scheme reflects
a common layer of representation that in a correct translation
is mostly shared between the translation and the source.

We note that verbal argument structures are less stable in this sense, as a verbal
argument structure may be translated or paraphrased to a nominal or adjectival
argument structure, and are thus less suitable as a basis for a semantic measure.
E.g., ``after he graduated, Tom moved to NYC'' contains two verbs,
but is similar in meaning to the one-verbed
``after graduation, Tom moved to NYC''.

@@

The Abstract Meaning Representation (AMR) scheme \cite{banarescu2013abstract}
shares much of UCCA's motivation for defining a more complete semantic annotation.
However, using AMR is not optimal for defining a decomposition of a sentence into semantic
units as it does not ground its semantic symbols in the text,
thus does not provide clear decomposition of the sentence into sub-units.
Also, AMR is more fine-grained than UCCA and consequently more difficult to annotate.
Other approaches pursued a grounded representation of semantics,
and have generally represented semantic structures as bi-lexical dependencies.
Examples include the tectogrammatical layer \parcite{sgallhp:1986} used in the Prague dependency
treebanks, e.g. the Czech-English one \parcite{hajic2012announcing}, as well
as dependencies derived from Minimal Recursion Semantics representations \parcite{oepen2006discriminant}.
Nevertheless, despite their appeal as deep and intuitive means for semantic representation,
these approaches require linguistic expertise for their annotation, due to the intricate and
fine-grained distinctions they provide, and are thus in our view less suitable for basing
an interpretable MT evaluation on.

@@

This section summarises the manual annotation procedure performed in order 
to compute the HUME measure. 
We mark the source sentence as $s$ and the translation sentence as $t$. 
The procedure involves two manual steps: (1) UCCA-annotating $s$, 
(2) human judgements as to the translation quality of each semantic unit of $s$ relative to $t$,
where units are defined according to the UCCA annotation.
We note that UCCA annotation is performed once for every source sentence,
irrespective of the number of its translations we wish
to evaluate\footnote{Instead of relying on the source,
the evaluation could be based on the UCCA structure for the \emph{reference}
translation. This would allow monolingual annotators to take part in the
evaluation. On the downside, the translation would have to be aligned to the
reference and more importantly, we would risk inaccurate judgements due to
the naturally occurring differences between the source and its reference
translations.}.

>>>>

\paragraph{Machine Translation Evaluation.}
%MT evaluation has been traditionally pursued through human-based
%evaluation that quantifies the adequacy of the translation in preserving the meaning 
%of the source sentence, and its fluency as a sentence in the target language \cite{lopez2008statistical}.
Human evaluation is generally done by ranking the outputs of multiple systems
e.g., in the WMT tasks \cite{bojar2015findings}, or by giving
adequacy/fluency scores to each translation, a procedure recently improved
by \newcite{graham2015improving}.
However, while providing the gold standard for MT evaluation, human evaluation
is not a scalable solution.
%and is infeasible where a wide variety of systems or parameter settings are to be evaluated.
% has motivated the construction of cheaper methods that correlate with human
%judgement.

Scalability is addressed by employing automatic and semi-automatic approximations of human
judgements. Commonly, such scores decompose over the sub-parts of
the translation, and quantify how many of these sub-parts appear in a manually created reference translation.
This decomposition allows system developers
to localize the errors, or at least unconfirmed parts within the translation.
The most commonly used measures decompose over n-grams or individual words, e.g., 
BLEU \cite{Papineni:2002}, NIST \cite{Doddington:2002} and METEOR  \cite{Banerjee:2005}.
Another common approach is to determine the similarity between the reference and translation
by applying string edit distance \cite{snover2006study}.
While these measures stimulated much progress in MT research by allowing
the evaluation of massive-scale experiments, the focus on words and n-grams does
not provide a good estimate of semantic correctness, and may favour shallow string-based
MT models.

@@

In order to address this shortcoming, more recent work quantified
the similarity of the reference and translation in terms
of their structure. \newcite{liu2005syntactic} took a syntactic approach, 
using dependency grammar, and
\newcite{owczarzak2007evaluating} took a similar approach using lexical-functional grammar structures.
\newcite{gimenez2007linguistic} proposed to combine multiple types of information,
capturing the overlap between the translation and reference in terms of their
semantic (predicate-argument structures), lexical and morphosyntactic features.

@@

{\small
\begin{tabular}{|l|l|l|l|}
\hline
L & Linker &  A & Participant \\
H & Scene & R & Relator \\
P & Process & C & Centre \\
\hline
\end{tabular}
}
    \end{center}
\caption{\label{fig:ucca_example_v2}
  Sample UCCA annotation where leaves correspond to words and nodes to units.
  Edge labels mark UCCA categories, listed  in the table.}
%which are largely ignored in this work
\end{figure}

\paragraph{Semantic Representation.}
UCCA (Universal Conceptual Cognitive Annotation, \inparcite{abend2013universal}) is a
cross-linguistically applicable, lightweight
scheme for semantic annotation. Formally, an UCCA structure is a directed acyclic graph (DAG),
whose leaves correspond to the words of the text.
The graph's nodes, called {\sc units}, are either terminals or several elements jointly
viewed as a single entity according to some semantic or cognitive consideration. Edges bear
a category, indicating the role of the sub-unit in the structure the unit
represents.

UCCA's current inventory of distinctions focuses on argument structures
(adjectival, nominal, verbal and others) and relations between them.
The most basic notion is the Scene, which describes a movement, an
action or a state which persists in time. Each Scene contains one main
relation and zero or more participants. For example, the sentence ``After graduation, Tom moved to NYC''
contains two Scenes, whose main relations are ``graduation'' and ``moved''.
The participant ``Tom'' is a part of both Scenes, while ``NYC'' only of the
latter (\figref{fig:ucca_example_v2}). Further categories account for
relations between Scenes and the internal structures of complex participants and relations.

% UCCA Motivation
UCCA is a natural candidate for defining a semantic
MT evaluation measure for a number of reasons. 
First, it defines an intuitive set of categories
that can be reliably annotated by non-experts, after as little as two hours
of training \cite{marinotti2014}.
Second, UCCA is a cross-linguistically applicable scheme, seeking to
represent what is shared between languages
by building on linguistic typological theory, and specifically on ``Basic Linguistic Theory''
\cite{Dixon:10a,Dixon:10b,Dixon:12}.
The cross-linguistic applicability of UCCA has been tested in annotations of
English, French, German and Czech.
Third, the scheme has been shown to be semantically stable
in translation: UCCA annotations of translated text usually contain the same set of relationships
\cite{sulem2015conceptual}. This indicates that the scheme reflects
a common layer of representation that in a correct translation
is mostly shared between the translation and the source.

We note that verbal argument structures are less stable in this sense, as a verbal
argument structure may be translated or paraphrased to a nominal or adjectival
argument structure, and is thus less suitable as a basis for a semantic measure.
E.g., ``after he graduated, Tom moved to NYC'' contains two verbs,
but is similar in meaning to the one-verbed
``after graduation, Tom moved to NYC''.

@@

The Abstract Meaning Representation (AMR) scheme \cite{banarescu2013abstract}
shares much of UCCA's motivation for defining a more complete semantic annotation.
However, using AMR is not optimal for defining a decomposition of a sentence into semantic
units as it does not ground its semantic symbols in the text,
thus does not provide clear decomposition of the sentence into sub-units.
Also, AMR is more fine-grained than UCCA and consequently harder to annotate.
Other approaches pursued a grounded representation of semantics,
and have generally represented semantic structures as bi-lexical dependencies.
Examples include the tectogrammatical layer \parcite{sgallhp:1986} used in the Prague dependency
treebanks, e.g. the Czech-English one \parcite{hajic2012announcing}, as well
as dependencies derived from Minimal Recursion Semantics representations \parcite{oepen2006discriminant}.
Nevertheless, despite their appeal as deep and intuitive means for semantic representation,
these approaches require linguistic expertise for their annotation, due to the intricate and
fine-grained distinctions they provide, and so we argue they are less suitable as a basis
for an interpretable MT evaluation.

@@

This section summarises the manual annotation procedure used
to compute the HUME measure. 
We mark the source sentence as $s$ and the translation as $t$. 
The procedure involves two manual steps: (1) UCCA-annotating $s$, 
(2) human judgements as to the translation quality of each semantic unit of $s$ relative to $t$,
where units are defined according to the UCCA annotation.
We note that UCCA annotation is performed once for every source sentence,
irrespective of the number of its translations we wish
to evaluate\footnote{Instead of relying on the source,
the evaluation could be based on the UCCA structure for the \emph{reference}
translation. This would allow monolingual annotators to take part in the
evaluation. On the downside, the translation would have to be aligned to the
reference and more importantly, we would risk inaccurate judgements due to
the naturally occurring differences between the source and its reference
translations.}.
